---
title: "Handling Rebounds with Probabilistic Constraints"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Handling Rebounds with Probabilistic Constraints}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(TConstrain)
library(igraph)
library(plotly)
```


# The Rebound Scenario

In partial reprogramming, cells often return to a state transcriptionally similar to their start point. 
Here, we simulate this scenario with 4 clusters:

    Day 0: Start

    Day 3: Transition (Reprogramming)

    Day 5: Reverting

    Day 10: Rebound (Similar to Day 0)
    
```{r}
# Mock Centroids
centroids_mock <- matrix(c(
  0, 0,  # Day 0
  5, 5,  # Day 3
  5, 0,  # Day 5
  1, 1   # Day 10 (Physically close to Day 0)
), ncol = 2, byrow = TRUE)
rownames(centroids_mock) <- c("Day0", "Day3", "Day5", "Day10")

# Simulate cluster labels for 100 cells
# (This step is just for the function signature, usually you have real data)
mock_dim <- centroids_mock[rep(1:4, each=25), ]
mock_labels <- rep(c("Day0", "Day3", "Day5", "Day10"), each=25)
```

# Defining Constraints

We enforce a "Must Link" between Day 0 and Day 3, and a "Cannot Link" between Day 0 and Day 10.

```{r}
cons <- data.frame(
  from = c("Day0", "Day0"),
  to = c("Day3", "Day10"),
  type = c("must_link", "cannot_link")
)

```

# Probabilistic Inference

We use probabilistic = TRUE to treat these as Bayesian priors. This allows the data to maintain some influence over the final topology.

```{r}
traj <- infer_trajectory(mock_dim, mock_labels, 
                         constraints = cons, 
                         probabilistic = TRUE, 
                         prior_strength = 20)

plot(traj$mst, layout=traj$centroids, main="Probabilistic Trajectory")
```


# Result

```{r}
# Visualize the resulting MST
plot(traj$mst, layout=traj$centroids, 
     main="Constrained Trajectory", 
     vertex.size=30, vertex.label.color="black")
```

The resulting edges correctly follow the temporal order: Day0 -> Day3 -> Day5 -> Day10, avoiding the shortcut between Day0 and Day10.
